---
title: "Schwannoma sample VS4 analysis"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(error = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_knit$set(root.dir = "/Users/ericpan/Documents/pan/usc/kgp/seurat/notebook")
knitr::opts_chunk$set(fig.align = "center")
```

Load packages needed and set system environment
```{r, include=FALSE}
library(dplyr)
library(Seurat)
library(patchwork)
library(tibble)
library(AnnotationHub)
library(limma)
library(edgeR)
library(ggplot2)
library(cowplot)
library(Matrix)
library(scales)
library(RCurl)
library(purrr)
library(metap)
library(multtest)
library(xlsx)
library(ggpubr)
library(clustree)
library(DESeq)
library(DESeq2)
library(ComplexHeatmap)
library(pheatmap)
library(reticulate)
library(SingleR)
library(scRNAseq)
library(scater)
library(DropletUtils)
library(stringr)
library(data.table)
library(janitor)
library(ggpubr)
library(rstatix)
library(RColorBrewer)
library(circlize)
use_python("/Users/ericpan/opt/miniconda3/bin/python3")
hpca.se <- HumanPrimaryCellAtlasData()
options(future.globals.maxSize = 4000 * 1024^2)
```

Get annotation information, gene name, entrez id and gene id
```{r}
ah <- AnnotationHub()
ahDb <- query(ah, 
              pattern = c("Homo sapiens", "EnsDb"), 
              ignore.case = TRUE)
id <- ahDb %>%
        mcols() %>%
        rownames() %>%
        tail(n = 1)
edb <- ah[[id]]
annotations <- genes(edb, 
                     return.type = "data.frame")
annotations <- annotations %>%
        dplyr::select(gene_id, gene_name, entrezid,description)
annotations$entrezid <- as.character(annotations$entrezid)

head(annotations, 10)
```

Read in 10x outputs direcotry and make seurat object (barcodes, features, matrix)
```{r}
VS4.data <- Read10X(data.dir = "/Users/ericpan/Documents/pan/usc/kgp/seurat/data/VS4_filtered_feature_bc_matrix")
VS4 <- CreateSeuratObject(counts = VS4.data, project = "VS4", min.features = 100)

VS4
```

QC and selecting cells for further analysis
```{r}
#Find mitochondrial genes
VS4[["percent.mt"]] <- PercentageFeatureSet(VS4, pattern = "^MT-")
#Visualize basic information of the dataset
VlnPlot(VS4, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

Remove outliers in the dataset
```{r}
VS4 <- subset(VS4, subset = nFeature_RNA > 200 & nFeature_RNA < 6000 & percent.mt < 20)
```

Gene level filtering
```{r}
# Gene level filtering
counts <- GetAssayData(object = VS4, slot = "counts")
nonzero <- counts > 0
keep_genes <- Matrix::rowSums(nonzero) >= 10
filtered_counts <- counts[keep_genes, ]
VS4 <- CreateSeuratObject(filtered_counts, meta.data = VS4@meta.data)
VS4
```

Perform SCTransform on the sample
```{r}
VS4 <- SCTransform(VS4, vars.to.regress = "percent.mt", verbose = FALSE)
# Note there is a new SCT assay now
VS4@assays
```


Perform dimensionality reduction by PCA and UMAP embedding
```{r, fig.align = "center"}
VS4 <- RunPCA(VS4, verbose = FALSE)
VS4 <- RunUMAP(VS4, dims = 1:30, umap.method = 'umap-learn', metric = 'correlation', verbose = F)

VS4 <- FindNeighbors(VS4, dims = 1:30, verbose = FALSE)
VS4 <- FindClusters(VS4, resolution = c(0,0.1,0.4,0.6,1.0), verbose = FALSE)
# Take a look at the clustering result
DimPlot(VS4, label = TRUE) + NoLegend()
```

Use Clustree function to see the cluster change while increasing the resolution
```{r, fig.align = "center", fig.width= 5, fig.asp = 1.4, eval=F}
clustree(VS4, prefix = "SCT_snn_res.")
```

Let's say we'd like to use 0.4 resolution to do the downstream analysis
```{r}
# Set the clustering resolution to 0.4
required_reso <- "SCT_snn_res.0.4"
Idents(object = VS4) <- required_reso
# Use SingleR package to do preliminary annotation on each cluster
DefaultAssay(VS4) <- "RNA"
VS4@meta.data$cells <- rownames(VS4@meta.data)
VS4.sc <- as.SingleCellExperiment(VS4)
colnames(VS4.sc) <- colData(VS4.sc)$cells
VS4.sc
```

Cluster level annotation using fine lables in singleR
```{r}
# cluster level annotation fine labels
pred.cluster_fine <- SingleR(test = VS4.sc, ref = hpca.se, labels = hpca.se$label.fine, method = "cluster", clusters = VS4@meta.data$SCT_snn_res.0.4)
cluster_anno <- pred.cluster_fine$labels
names(cluster_anno) <- paste0("cluster",c(0:(length(cluster_anno)-1)))
cluster_anno <- as.data.frame(cluster_anno)
head(cluster_anno)
```

Cell level annotation
```{r}
# Get all cell ids in a cluster
cluster_barcode <- lapply(levels(VS4@meta.data$SCT_snn_res.0.4), function(i) VS4@meta.data[which(VS4@meta.data$SCT_snn_res.0.4 == i),"cells"])
pred.cell_fine <- sapply(cluster_barcode, function(i) VS4.sc[,i] %>% SingleR(ref = hpca.se, labels = hpca.se$label.fine))
cell_fine_labels <- sapply(pred.cell_fine, function(i) table(i$labels) %>% as.data.frame() %>% 
                        arrange(desc(Freq)) %>% pull(1) %>% as.character() %>% head(3)) %>% as.data.frame()
cell_fine_labels <- t(cell_fine_labels) %>% as.data.frame()
rownames(cell_fine_labels) <- paste0("cluster", c(0:(length(cluster_barcode)-1)))
cell_fine_labels[1:5,]
```

Find cluster marker genes
```{r}
Idents(VS4) <- "SCT_snn_res.0.4"
all_markers <- FindAllMarkers(object = VS4, 
                          only.pos = TRUE,
                          logfc.threshold = 0.20,
                          verbose = FALSE)
top30_markers <- all_markers %>% group_by(cluster) %>% top_n(n = 30, wt = avg_logFC)
top30_markers <- left_join(x = top30_markers, y = unique(annotations[, c("gene_name", "description","entrezid")]),
                                 by = c("gene" = "gene_name")) %>% as.data.frame()
top30_markers[1:5,]
```


